Something about Performance, 2D Component of NEMO. (TBC)

### 2015-06-08 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
Add text describing nemolite2d provided by Hedong

\author{J.~Appleyard, M.~Ashworth, R.~W.~Ford, J.~Holt, H.~Lieu, A.~R.~Porter and G.~D.~Riley% <-this % stops a space
\thanks{J.~Appleyard is with NVIDIA.}%
\thanks{H.~Lieu and J.~Holt are with the National Oceanography Centre, Liverpool, UK.}%
\thanks{M.~Ashworth, R.~W.~Ford and A.~R.~Porter are with the Science and Technology Facilities Council, UK.}% <-this % stops a space
\thanks{G.~Riley is with the University of Manchester, Manchester, UK.}% <-this % stops a space
}

@@

\subsection{The `nemolite2d' Program}
For this work we have used a  program

{\bf Description of the structure of the program at the Algorithm
  level so that it links in to Profile data in results section.}

@@

As with the serial optimisations, we consider the effect of each of
the OpenMP optimisation steps described in Section~\ref{sec_omp_steps}
for the $256^2$ domain. For this we principally use a single Intel,
Ivy Bridge socket which has 12 hardware cores and support for up to 24
threads with hyperthreading.
Figures~\ref{FIG_gnu_omp_256},~\ref{FIG_intel_omp_256}
and~\ref{FIG_cray_omp_256} show the performance of each of the
versions of the code on this system for the Gnu, Intel and Cray
compilers, respectively.

@@

@@

\begin{figure}
\centering
\includegraphics[width=\picwidth]{omp_scaling_32_gnu}
\caption{Performance of the OpenMP-parallel version of \psykal
  nemolite2d for the Gnu compiler on a single Intel Ivy Bridge
  socket. The corresponding parallel efficiencies are shown using open
  symbols and dashed lines. The 24-thread runs employed hyperthreading
  and the optimal OpenMP schedule is given in parentheses.}
\label{FIG_gnu_omp_32}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\picwidth]{omp_scaling_32_intel}
\caption{Performance of the OpenMP-parallel version of \psykal
  nemolite2d for the Intel compiler on a single Intel Ivy Bridge
  socket.  The corresponding parallel efficiencies are shown using
  open symbols and dashed lines. The 24-thread runs employed
  hyperthreading and the optimal OpenMP schedule is given in
  parentheses.}
\label{FIG_intel_omp_32}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\picwidth]{omp_scaling_32_cray}
\caption{Performance of the OpenMP-parallel version of \psykal
  nemolite2d for the Cray compiler on a single Intel Ivy Bridge
  socket.  The corresponding parallel efficiencies are shown using
  open symbols and dashed lines. The 24-thread runs employed
  hyperthreading and the optimal OpenMP schedule is given in
  parentheses.}
\label{FIG_cray_omp_32}
\end{figure}

>>>>

\author{J.~Appleyard, M.~Ashworth, R.~W.~Ford, J.~Holt, H.~Liu, A.~R.~Porter and G.~D.~Riley% <-this % stops a space
\thanks{J.~Appleyard is with NVIDIA.}%
\thanks{H.~Liu and J.~Holt are with the National Oceanography Centre, Liverpool, UK.}%
\thanks{M.~Ashworth, R.~W.~Ford and A.~R.~Porter are with the Science and Technology Facilities Council, UK.}% <-this % stops a space
\thanks{G.~Riley is with the University of Manchester, Manchester, UK.}% <-this % stops a space
}

@@

\subsection{The `nemolite2d' Program}

For this work we have used a program "nemolite2d" developed by
ourselves.  “nemolite2d” is a vertically-averaged version of NEMO
(Nucleus for European Modelling of the Ocean), which only keeps its
dynamical part. The whole model system is represented through one
continuity equation (for the update of the sea-surface height) and two
vertically-integrated momentum equations (for the two velocity
components, respectively).

(add above three equations in .tex file).

The external forcing includes surface wind stress, bottom friction,
and open boundary barotropic forcing. A lateral slip boundary
condition is applied along the coast lines. The open boundary
conditions can be set as a clipped or a Flather's radiation
condition (Flather, 1976). The bottom friction takes a semi-implicit
form for the sake of model stability. As done in the original version
of NEMO, a constant or Smagorinsky horizontal viscosity coefficient is
used for the horizontal viscosity term.

The traditional Arakawa C structured grid is employed here for the
discretisation of the computational domain. A two-dimensional integer
array is used to identify the different parts of the computational
domain; it has the value of one for ocean, zero for land and minus
one for ocean cells outside of the computational domain. This array
serves as a mark to identify ocean cells, land cells, solid boundaries
and open boundaries.

For the sake of simplicity, the explicit Eulerian forward time
stepping method is implemented here, except that the bottom friction
takes a semi-implicit form.  The Coriolis force can be set in explicit
or implicit form. The advection term is computed with a first-order
upwind scheme.

The order of the model computation is as follows:
\begin{enumerate}
\item Set the initial conditions (water depth, sea surface height, velocity)
\item integrate the continuity equation for the new sea surface height
\item update the different terms in the right hand side of the momentum equations (advection, Coriolis forcing(if set in explicit form), pressure gradient, and, horizontal viscosity
\item update the velocity vectors by summing up the values in 3), and implicitly with the bottom friction and Coriolis forcing (if set in implicit form)
\item apply the boundary conditions on to the open boundary and solid boundary cells.
\end{enumerate}

@@

As with the serial optimisations, we consider the effect of each of
the OpenMP optimisation steps described in Section~\ref{sec_omp_steps}
for the $256^2$ domain. For this we principally use a single Intel,
Ivy Bridge socket which has 12 hardware cores and support for up to 24
threads with hyperthreading (i.e. two threads per core).
Figures~\ref{FIG_gnu_omp_256},~\ref{FIG_intel_omp_256}
and~\ref{FIG_cray_omp_256} show the performance of each of the
versions of the code on this system for the Gnu, Intel and Cray
compilers, respectively.

@@

Since the space to explore consists of three different compilers, six
different domain sizes, five stages of optimisation and six different
thread counts we can only consider a slice through it in what follows.
In order to inform our choice of domain size,
Figure~\ref{FIG_omp_domain_size} shows the scaling behaviour of the
most performant, Intel-compiled version of nemolite2d. Although it is
common for production runs of NEMO to use MPI sub-domains of $20
\times 20$ ({\bf do we have/need a reference for that?}), it is clear
from Figure~\ref{FIG_omp_domain_size} that the limited quantity of
parallelism in the $32 \times 32$ domain inhibits scaling. Therefore,
to fully test the performance of our OpenMP implementations we have
chosen to examine results for the $256 \times 256$ domain and these
are shown in Figures~\ref{FIG_gnu_omp_256}, \ref{FIG_intel_omp_256}
and \ref{FIG_cray_omp_256}.

\begin{figure}
\centering
\includegraphics[width=\picwidth]{omp_scaling_problem_size}
\caption{The scaling behaviour of the most performant OpenMP-parallel
  version of \psykal nemolite2d for the full range of domain sizes
  considered on the CPU. Results are for the Intel compiler on a
  single Intel Ivy Bridge socket. The corresponding parallel
  efficiencies are shown using open symbols and dashed lines. The
  24-thread runs employed hyperthreading.}
\label{FIG_omp_domain_size}
\end{figure}

@@

%\begin{figure}
%\centering
%\includegraphics[width=\picwidth]{omp_scaling_32_gnu}
%\caption{Performance of the OpenMP-parallel version of \psykal
%  nemolite2d for the Gnu compiler on a single Intel Ivy Bridge
%  socket. The corresponding parallel efficiencies are shown using open
%  symbols and dashed lines. The 24-thread runs employed hyperthreading
%  and the optimal OpenMP schedule is given in parentheses.}
%\label{FIG_gnu_omp_32}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=\picwidth]{omp_scaling_32_intel}
%\caption{Performance of the OpenMP-parallel version of \psykal
%  nemolite2d for the Intel compiler on a single Intel Ivy Bridge
%  socket.  The corresponding parallel efficiencies are shown using
%  open symbols and dashed lines. The 24-thread runs employed
%  hyperthreading and the optimal OpenMP schedule is given in
%  parentheses.}
%\label{FIG_intel_omp_32}
%\end{figure}

%\begin{figure}
%\centering
%\includegraphics[width=\picwidth]{omp_scaling_32_cray}
%\caption{Performance of the OpenMP-parallel version of \psykal
%  nemolite2d for the Cray compiler on a single Intel Ivy Bridge
%  socket.  The corresponding parallel efficiencies are shown using
%  open symbols and dashed lines. The 24-thread runs employed
%  hyperthreading and the optimal OpenMP schedule is given in
%  parentheses.}
%\label{FIG_cray_omp_32}
%\end{figure}
